On type cones of $\b{g}$-vector fans

### 2020-04-03 thirdPaper/FPSAC/typeConeAssociahedraFPSAC.tex

Fix an essential complete simplicial fan\footnote{A \defn{polyhedral cone} is the positive span of finitely many vectors or equivalently, the intersection of finitely many closed linear half-spaces. The \defn{faces} of a cone are its intersections with its supporting hyperplanes. A \defn{fan}~$\Fan$ is a set of polyhedral cones such that any face of a cone of~$\Fan$ belongs to~$\Fan$, and any two cones of~$\Fan$ intersect along a face of both. A fan is \defn{essential} if the intersection of its cones is the origin, \defn{complete} if the union of its cones covers~$\R^n$, and \defn{simplicial} if all its cones are generated by dimension many rays.}~$\Fan$ in~$\R^n$. Let~$\b{G}$ be the $N \times n$-matrix whose rows are (representative vectors of) the rays of~$\Fan$ and let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$). For any height vector~$\b{h} \in \R^N$, we define the polytope\footnote{A \defn{polytope} is the convex hull of finitely many points or equivalently, a bounded intersection of finitely many closed affine half-spaces.}
\(
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}.
\)
Unfortunately, the normal fan\footnote{The \defn{normal cone} of a face~$F$ of a polytope~$P$ is the cone generated by the normal vectors of the facets of~$P$ containing~$F$. The \defn{normal fan} of~$P$ is the set of normal cones of all its faces.} of this polytope~$P_\b{h}$ does not always coincide with the fan~$\Fan$ we started from. The following statement gives a simple characterization of the height vectors~$h$ for which it is the case. It is a reformulation of regularity of triangulations of vector configurations, introduced in the theory of secondary polytopes~\cite{GelfandKapranovZelevinsky}.
We present here a convenient formulation from~\cite{ChapotonFominZelevinsky}.

@@

Note that the type cone is an open cone and contains a linearity subspace of dimension~$n$ (it is invariant by translation in~$\b{G} \R^n$). Therefore, we say that the type cone is \defn{simplicial} when it has precisely $N-n$ facets.

\begin{definition}
An \defn{extremal adjacent pair} of~$\Fan$ is a pair of adjacent maximal cones~$\{\R_{\ge0}\rays, \R_{\ge0}\rays'\}$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0$ in the definition of the type cone~$\typeCone(\Fan)$ actually defines a facet of~$\typeCone(\Fan)$.
In other words, extremal adjacent pairs define the extremal rays of the polar of the closed type cone~$\typeCone(\Fan)$.
\vspace{.2cm}
\end{definition}

\begin{example}
\label{exm:typeCone}
\enlargethispage{.1cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left).
% (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional linearity space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{xy}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$x$ and~$y$.
%Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant, while the other three define facets, so that the extremal adjacent pairs are~$(a,b)$, $(b,c)$ and~$(c,d)$.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, we slice it with a transversal $2$-dimensional affine space. This is the red triangle depicted in \cref{fig:typeCone} (right).

@@

\begin{remark}
All fans considered in \cref{sec:gvectorFans} have the \defn{unique exchange relation property}: the linear dependence ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0$ only depends on the exchanged rays~$\ray[r]$ and~$\ray[r]'$, not on the maximal adjacent cones~$\rays$ and~$\rays'$. We then write~$\coefficient[{\ray[s]}][\ray][\ray']$ instead of~$\coefficient[{\ray[s]}][\rays][\rays']$ and we call \defn{extremal exchangeable pairs} of~$\Fan$ the pairs of exchangeable rays~$\{\ray, \ray'\}$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\ray][\ray'] \, \b{h}_{\ray[s]} > 0$ defines a facet of~$\typeCone(\Fan)$.

@@

To conclude this section, we consider alternative polytopal realizations of the fan~$\Fan$ and discuss their behavior in the situation when the type cone~$\typeCone(\Fan)$ is simplicial.

@@

%We start with the classical associahedron.
An $n$-dimensional \defn{associahedron} is a polytope whose face lattice is the reverse inclusion lattice of dissections (\ie sets of pairwise non-crossing diagonals) of a convex~$(n+3)$-gon.
In particular, its vertices correspond to triangulations of the $(n+3)$-gon and its facets correspond to internal diagonals of the $(n+3)$-gon.
Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ (resp.~$Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n}$) denote all (resp.~the internal) diagonals.
The \defn{$\b{g}$-vector} of a diagonal $(a,b) \in X(n)$ is ${\gvector{a,b} \eqdef {\displaystyle\sum}_{\,a < \ell < b} \b{e}_\ell - \frac{b-a-1}{n+1} {\displaystyle\sum}_{\,1 \le \ell \le n+1} \b{e}_\ell}$.
%Note that~$\gvector{a,b} = 0$ if~$(a,b)$ is not internal.
We set~$\gvector{D} \eqdef \set{\gvector{a,b}}{(a,b) \in D}$ for a dissection~$D$.
Recall that:
\begin{compactitem}
\item the set of cones
\(
\gvectorFan[n] \eqdef \bigset{\R_{\ge0} \, \gvectors{D}}{D \text{ dissection of the $(n+3)$-gon}}
\)
forms a complete simplicial fan.
See \cref{fig:lodayFans}\,(left and middle).

\item the fan~$\gvectorFan[n]$ is the normal fan of an associahedron~$\Asso[n]$ constructed \eg in~\cite{Loday}.

@@

\begin{corollary}
\label{coro:typeConeAsso}
%Let~$n \in \N$ and~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$. 
For any~$n \in \N$, the type cone of the fan~$\gvectorFan[n]$ is given by
\[
\typeCone \big( \gvectorFan[n] \big) = \set{\b{h} \in \R^{X(n)}}{\begin{array}{@{}l@{}} \b{h}_{(0,n+2)} = 0, \qquad\text{and}\qquad \b{h}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{h}_{(a,b)} \! + \! \b{h}_{(a',b')} \! > \! \b{h}_{(a,b')} \! + \! \b{h}_{(a',b)} \text{ for all } 0 \! \le \! a \! < \! a' \! < \! b \! < \! b' \! \le \! n+2 \end{array}} \!.
\]
\end{corollary}

@@

%\begin{example}
%\label{exm:typeConeAsso}
%The type cone of~$\gvectorFan[2]$ (resp.~$\gvectorFan[3]$) lives in~$\R^5$ (resp.~$\R^9$), has a linearity space of dimension~$2$ (resp.~$3$), and has $3$ (resp.~$6$) facet-defining inequalities described in \cref{prop:extremalExchangeablePairsAsso}.
%We represent these inequalities in red in \cref{fig:labelFacetDefiningInequalititiesAsso}, using incoming and outgoing arrows to distinguish their left and right hand sides.
%For instance, the inequality~$\red \circled{A}$ is given by~$\b{h}_{\includegraphics[scale=.3]{diagonalA1}} + \b{h}_{\includegraphics[scale=.3]{diagonalA3}} > \b{h}_{\includegraphics[scale=.3]{diagonalA2}}$.

@@

\begin{corollary}[{\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}}]
\label{coro:allPolytopalRealizationsAsso}
%Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ and $Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n}$.
For any~$n \in \N$ and any~$\b{\ell} \in \R_{>0}^{Y(n)}$, the polytope
\[
R_\b{\ell}(n) \eqdef \set{\b{z} \in \R^{X(n)}}{\begin{array}{@{}l@{}} \b{z} \ge 0, \quad \b{z}_{(0,n+2)} = 0 \quad\text{and}\quad \b{z}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{z}_{(a,b-1)} + \b{z}_{(a+1,b)} - \b{z}_{(a,b)} - \b{z}_{(a+1,b-1)} = \b{\ell}_{(a,b)} \text{ for all } (a,b) \in Y(n) \end{array}}
\]
is an $n$-dimensional associahedron, whose normal fan is~$\gvectorFan[n]$.
Moreover, the polytopes~$R_\b{\ell}(n)$ for~$\b{\ell} \in \R_{>0}^{Y(n)}$ describe all polytopal realizations of the fan~$\gvectorFan[n]$.
\end{corollary}

@@

\item the cluster fan is the normal fan of a polytope~$\Asso[\B_\circ]$, called \defn{generalized associa}\-\defn{hedron}, constructed for bipartite (resp.~acyclic, resp.~arbitrary) initial exchange matrices in~\cite{ChapotonFominZelevinsky} (resp.~\cite{HohlwegLangeThomas}, resp.~\cite{HohlwegPilaudStella}).

@@

%Its type cone lives in~$\R^{9}$ and has a linearity space of dimension~$3$.

@@

\begin{figure}[b]
	\newcommand{\fitClusterVariable}[1]{\parbox[c][.7cm][c]{1.5cm}{\centering $#1$}}
	\centerline{
    \begin{tikzpicture}
    	\matrix (m) [matrix of math nodes, row sep=.4cm, column sep=-.1cm, nodes={anchor=center, align=center, inner sep=0pt}, ampersand replacement=\&]{
    		\fitClusterVariable{\phantom{1}} \& \node (e1) {\red \circled{E}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_2 x_3}} \&\& \fitClusterVariable{x_3} \& \node (c) {\red \circled{C}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_1 x_3}} \&\& \fitClusterVariable{x_1} \& \node (a2) {\red \circled{A}}; \& \fitClusterVariable{\phantom{1}} \\
    		\& \fitClusterVariable{\frac{x_1 + x_2}{x_3}} \& \node (f1) {\red \circled{F}}; \& \fitClusterVariable{\frac{x_1 + x_3}{x_2}} \& \node (b) {\red \circled{B}}; \& \fitClusterVariable{\frac{x_2 + x_3}{x_1}} \& \node (d) {\red \circled{D}}; \& \fitClusterVariable{\frac{x_1 + x_2}{x_3}} \& \node (f2) {\red \circled{F}}; \& \fitClusterVariable{\frac{x_1 + x_3}{x_2}} \\
    		\fitClusterVariable{\phantom{1}} \&\& \fitClusterVariable{x_1} \& \node (a1) {\red \circled{A}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_1 x_2}} \&\& \fitClusterVariable{x_2} \& \node (e2) {\red \circled{E}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_2 x_3}} \&\& \fitClusterVariable{\phantom{1}} \\};
    	\draw[densely dotted, thick] (m-1-1) -- (m-2-2);
    	\draw[densely dotted, thick] (m-3-1) -- (m-2-2);
    	\draw[->] (m-2-2) -- (m-1-3);
    	\draw[->] (m-2-2) -- (m-3-3);
    	\draw[->] (m-1-3) -- (m-2-4);
    	\draw[->] (m-3-3) -- (m-2-4);
    	\draw[->] (m-2-4) -- (m-1-5);
    	\draw[->] (m-2-4) -- (m-3-5);
    	\draw[->] (m-1-5) -- (m-2-6);
    	\draw[->] (m-3-5) -- (m-2-6);
    	\draw[->] (m-2-6) -- (m-1-7);
    	\draw[->] (m-2-6) -- (m-3-7);
    	\draw[->] (m-1-7) -- (m-2-8);
    	\draw[->] (m-3-7) -- (m-2-8);
    	\draw[->] (m-2-8) -- (m-1-9);
    	\draw[->] (m-2-8) -- (m-3-9);
    	\draw[->] (m-1-9) -- (m-2-10);
    	\draw[->] (m-3-9) -- (m-2-10);
    	\draw[densely dotted, thick] (m-2-10) -- (m-1-11);
    	\draw[densely dotted, thick] (m-2-10) -- (m-3-11);
    	\draw[red, -<] (a1) -- (m-3-3);
    	\draw[red, -<] (a1) -- (m-3-5);
    	\draw[red, ->] (a1) -- (m-2-4);
    	\draw[red, -<] (a2) -- (m-1-9);
    	\draw[red, densely dotted, thick] (a2) -- (m-1-11);
    	\draw[red, ->] (a2) -- (m-2-10);
    	\draw[red, -<] (b) -- (m-2-4);
    	\draw[red, -<] (b) -- (m-2-6);
    	\draw[red, ->] (b) -- (m-1-5);
    	\draw[red, ->] (b) -- (m-3-5);
    	\draw[red, -<] (c) -- (m-1-5);
    	\draw[red, -<] (c) -- (m-1-7);
    	\draw[red, ->] (c) -- (m-2-6);
    	\draw[red, -<] (d) -- (m-2-6);
    	\draw[red, -<] (d) -- (m-2-8);
    	\draw[red, ->] (d) -- (m-1-7);
    	\draw[red, ->] (d) -- (m-3-7);
    	\draw[red, densely dotted, thick] (e1) -- (m-1-1);
    	\draw[red, -<] (e1) -- (m-1-3);
    	\draw[red, ->] (e1) -- (m-2-2);
    	\draw[red, -<] (e2) -- (m-3-7);
    	\draw[red, -<] (e2) -- (m-3-9);
    	\draw[red, ->] (e2) -- (m-2-8);
    	\draw[red, -<] (f1) -- (m-2-2);
    	\draw[red, -<] (f1) -- (m-2-4);
    	\draw[red, ->] (f1) -- (m-1-3);
    	\draw[red, ->] (f1) -- (m-3-3);
    	\draw[red, -<] (f2) -- (m-2-8);
    	\draw[red, -<] (f2) -- (m-2-10);
    	\draw[red, ->] (f2) -- (m-1-9);
    	\draw[red, ->] (f2) -- (m-3-9);
    \end{tikzpicture}
    }
    % \vspace{-.1cm}
	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[\B_\circ] \big)$ for the cluster fan of \cref{fig:lodayFans}\,(right). See \cref{thm:extremalExchangeablePairsCA}. The conventions are similar to \cref{fig:labelFacetDefiningInequalititiesAsso}.}
	\label{fig:labelFacetDefiningInequalititiesCA}
\end{figure}

@@

\begin{theorem}
\label{thm:allPolytopalRealizationsCA}
For any finite type exchange matrix~$\B_\circ$, and for any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
\b{R}_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\variables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \!\!\!\!\! \sum_{y \in \variables} \!\!\!\!\!\coefficient[y][x][x'] \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$\b{R}_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{theorem}

@@

\item the non-kissing fan is the normal fan of a polytope~$\Asso[\quiver]$, called \defn{gentle associahedron}, constructed in~\cite{PaluPilaudPlamondon-nonkissing}.

>>>>

Fix an essential complete simplicial fan\footnote{A \defn{polyhedral cone} is the positive span of finitely many vectors or equivalently, the intersection of finitely many closed linear half-spaces. The \defn{faces} of a cone are its intersections with its supporting hyperplanes. A \defn{fan}~$\Fan$ is a set of polyhedral cones such that any face of a cone of~$\Fan$ belongs to~$\Fan$, and any two cones of~$\Fan$ intersect along a face of both. A fan is \defn{essential} if the intersection of its cones is the origin, \defn{complete} if the union of its cones covers~$\R^n$, and \defn{simplicial} if all its cones are generated by dimension many rays.}~$\Fan$ in~$\R^n$. Let~$\b{G}$ be the $N \times n$-matrix whose rows are (representative vectors of) the rays of~$\Fan$.
% and let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$). 
For any height vector~$\b{h} \in \R^N$, we define the polytope\footnote{A \defn{polytope} is the convex hull of finitely many points or equivalently, a bounded intersection of finitely many closed affine half-spaces.}
\(
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}.
\)
Unfortunately, the normal fan\footnote{The \defn{normal cone} of a face~$F$ of a polytope~$P$ is the cone generated by the normal vectors of the facets of~$P$ containing~$F$. The \defn{normal fan} of~$P$ is the set of normal cones of all its faces.} of this polytope~$P_\b{h}$ does not always coincide with the fan~$\Fan$ we started from. The following statement gives a simple characterization of the height vectors~$h$ for which it is the case. It is a reformulation of regularity of triangulations of vector configurations, introduced in the theory of secondary polytopes~\cite{GelfandKapranovZelevinsky}.
We present here a convenient formulation from~\cite{ChapotonFominZelevinsky}.

@@

Note that the type cone is an open cone and contains a lineality subspace of dimension~$n$ (it is invariant by translation in~$\b{G} \R^n$). Therefore, we say that the type cone is \defn{simplicial} when it has precisely $N-n$ facets.

\begin{definition}
An \defn{extremal adjacent pair} of~$\Fan$ is a pair of adjacent maximal cones~$\R_{\ge0}\rays, \R_{\ge0}\rays'$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0$ in the definition of the type cone~$\typeCone(\Fan)$ actually defines a facet of~$\typeCone(\Fan)$.
In other words, extremal adjacent pairs define the extremal rays of the polar of the closed type cone~$\typeCone(\Fan)$.
\vspace{.2cm}
\end{definition}

\begin{example}
\label{exm:typeCone}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left).
% (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional lineality space, and is defined by the following five inequalities:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{xy}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$x$ and~$y$.
%Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant, while the other three define facets, so that the extremal adjacent pairs are~$(a,b)$, $(b,c)$ and~$(c,d)$.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent $\typeCone(\Fan)$, we slice it with a transversal $2$-dimensional affine space, and obtain the red triangle in \cref{fig:typeCone} (right).

@@

\begin{remark}
All fans considered in \cref{sec:gvectorFans} have the \defn{unique exchange relation property}: the linear dependence ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0$ only depends on the exchanged rays~$\ray[r]$ and~$\ray[r]'$, not on the maximal adjacent cones~$\rays$ and~$\rays'$. We then write~$\coefficient[{\ray[s]}][\ray][\ray']$ instead of~$\coefficient[{\ray[s]}][\rays][\rays']$ and we call \defn{extremal exchangeable pairs} of~$\Fan$ the pairs of exchangeable rays~$\ray, \ray'$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\ray][\ray'] \, \b{h}_{\ray[s]} > 0$ defines a facet of~$\typeCone(\Fan)$.

@@

To conclude this section, we consider alternative polytopal realizations of the fan~$\Fan$ and discuss their behavior in the situation when the type cone~$\typeCone(\Fan)$ is simplicial.
Fix a $(N-n) \times N$-matrix~$\b{K}$ that spans the left kernel of the $N \times n$-matrix~$\b{G}$ (\ie $\b{K}\b{G} = 0$).

@@

%We start with the classical associahedron.
An $n$-dimensional \defn{associahedron} is a polytope whose face lattice is the reverse inclusion lattice of dissections (\ie sets of pairwise non-crossing diagonals) of a convex~$(n+3)$-gon.
In particular, its vertices correspond to triangulations of the $(n+3)$-gon and its facets correspond to internal diagonals of the $(n+3)$-gon.
Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ denote all diagonals of the $(n+3)$-gon and~$Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n} \subset X(n)$.
The \defn{$\b{g}$-vector} of a diagonal $(a,b) \in X(n)$ is ${\gvector{a,b} \eqdef {\displaystyle\sum}_{\,a < \ell < b} \b{e}_\ell - \frac{b-a-1}{n+1} {\displaystyle\sum}_{\,1 \le \ell \le n+1} \b{e}_\ell}$.
%Note that~$\gvector{a,b} = 0$ if~$(a,b)$ is not internal.
We set~$\gvector{D} \eqdef \set{\gvector{a,b}}{(a,b) \in D}$ for a dissection~$D$.
Recall that:
\begin{compactitem}
\item the set of cones
\(
\gvectorFan[n] \eqdef \bigset{\R_{\ge0} \, \gvectors{D}}{D \text{ dissection of the $(n+3)$-gon}}
\)
forms a complete simplicial fan.
See \cref{fig:lodayFans}\,(left and middle).

\item the fan~$\gvectorFan[n]$ is the normal fan of the associahedron~$\Asso[n]$ constructed \eg in~\cite{Loday}.

@@

\begin{corollary}
\label{coro:typeConeAsso}
%Let~$n \in \N$ and~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$. 
For any~$n \in \N$, the type cone of the fan~$\gvectorFan[n]$ is given by
\[
\typeCone \big( \gvectorFan[n] \big) = \set{\b{h} \in \R^{X(n)} \!}{\begin{array}{@{}l@{}} \b{h}_{(0,n+2)} = 0, \qquad\text{and}\qquad \b{h}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{h}_{(a,b)} \! + \! \b{h}_{(a',b')} \! > \! \b{h}_{(a,b')} \! + \! \b{h}_{(a',b)} \text{ for all } 0 \! \le \! a \! < \! a' \! < \! b \! < \! b' \! \le \! n+2 \end{array}} \!.
\]
\end{corollary}

@@

%\begin{example}
%\label{exm:typeConeAsso}
%The type cone of~$\gvectorFan[2]$ (resp.~$\gvectorFan[3]$) lives in~$\R^5$ (resp.~$\R^9$), has a lineality space of dimension~$2$ (resp.~$3$), and has $3$ (resp.~$6$) facet-defining inequalities described in \cref{prop:extremalExchangeablePairsAsso}.
%We represent these inequalities in red in \cref{fig:labelFacetDefiningInequalititiesAsso}, using incoming and outgoing arrows to distinguish their left and right hand sides.
%For instance, the inequality~$\red \circled{A}$ is given by~$\b{h}_{\includegraphics[scale=.3]{diagonalA1}} + \b{h}_{\includegraphics[scale=.3]{diagonalA3}} > \b{h}_{\includegraphics[scale=.3]{diagonalA2}}$.

@@

\begin{corollary}[{\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}}]
\label{coro:allPolytopalRealizationsAsso}
%Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ and $Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n}$.
For any~$n \in \N$ and any~$\b{\ell} \in \R_{>0}^{Y(n)}$, the polytope
\[
R_\b{\ell}(n) \eqdef \set{\b{z} \in \R^{X(n)} \!}{\begin{array}{@{}l@{}} \b{z} \ge 0, \quad \b{z}_{(0,n+2)} = 0 \quad\text{and}\quad \b{z}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{z}_{(a,b-1)} + \b{z}_{(a+1,b)} - \b{z}_{(a,b)} - \b{z}_{(a+1,b-1)} = \b{\ell}_{(a,b)} \text{ for all } (a,b) \in Y(n) \end{array}}
\]
is an $n$-dimensional associahedron, whose normal fan is~$\gvectorFan[n]$.
Moreover, the polytopes~$R_\b{\ell}(n)$ for~$\b{\ell} \in \R_{>0}^{Y(n)}$ describe all polytopal realizations of the fan~$\gvectorFan[n]$.
\end{corollary}

@@

\item the cluster fan~$\gvectorFan[\B_\circ]$ is the normal fan of the \defn{generalized associahedron}~$\Asso[\B_\circ]$, constructed for bipartite (resp.~acyclic, resp.~arbitrary) initial exchange matrices in~\cite{ChapotonFominZelevinsky} (resp.~\cite{HohlwegLangeThomas}, resp.~\cite{HohlwegPilaudStella}).

@@

%Its type cone lives in~$\R^{9}$ and has a lineality space of dimension~$3$.

@@

\begin{figure}[b]
	\newcommand{\fitClusterVariable}[1]{\parbox[c][.7cm][c]{1.6cm}{\centering $#1$}}
	\centerline{
    \begin{tikzpicture}
    	\matrix (m) [matrix of math nodes, row sep=.4cm, column sep=-.1cm, nodes={anchor=center, align=center, inner sep=0pt}, ampersand replacement=\&]{
    		\fitClusterVariable{\phantom{1}} \& \node (e1) {\red \circled{E}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_2 x_3}} \&\& \fitClusterVariable{x_3} \& \node (c) {\red \circled{C}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_1 x_3}} \&\& \fitClusterVariable{x_1} \& \node (a2) {\red \circled{A}}; \& \fitClusterVariable{\phantom{1}} \\
    		\& \fitClusterVariable{\frac{x_1 + x_2}{x_3}} \& \node (f1) {\red \circled{F}}; \& \fitClusterVariable{\frac{x_1 + x_3}{x_2}} \& \node (b) {\red \circled{B}}; \& \fitClusterVariable{\frac{x_2 + x_3}{x_1}} \& \node (d) {\red \circled{D}}; \& \fitClusterVariable{\frac{x_1 + x_2}{x_3}} \& \node (f2) {\red \circled{F}}; \& \fitClusterVariable{\frac{x_1 + x_3}{x_2}} \\
    		\fitClusterVariable{\phantom{1}} \&\& \fitClusterVariable{x_1} \& \node (a1) {\red \circled{A}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_1 x_2}} \&\& \fitClusterVariable{x_2} \& \node (e2) {\red \circled{E}}; \& \fitClusterVariable{\frac{x_1 + x_2 + x_3}{x_2 x_3}} \&\& \fitClusterVariable{\phantom{1}} \\};
    	\draw[densely dotted, thick] (m-1-1) -- (m-2-2);
    	\draw[densely dotted, thick] (m-3-1) -- (m-2-2);
    	\draw[->] (m-2-2) -- (m-1-3);
    	\draw[->] (m-2-2) -- (m-3-3);
    	\draw[->] (m-1-3) -- (m-2-4);
    	\draw[->] (m-3-3) -- (m-2-4);
    	\draw[->] (m-2-4) -- (m-1-5);
    	\draw[->] (m-2-4) -- (m-3-5);
    	\draw[->] (m-1-5) -- (m-2-6);
    	\draw[->] (m-3-5) -- (m-2-6);
    	\draw[->] (m-2-6) -- (m-1-7);
    	\draw[->] (m-2-6) -- (m-3-7);
    	\draw[->] (m-1-7) -- (m-2-8);
    	\draw[->] (m-3-7) -- (m-2-8);
    	\draw[->] (m-2-8) -- (m-1-9);
    	\draw[->] (m-2-8) -- (m-3-9);
    	\draw[->] (m-1-9) -- (m-2-10);
    	\draw[->] (m-3-9) -- (m-2-10);
    	\draw[densely dotted, thick] (m-2-10) -- (m-1-11);
    	\draw[densely dotted, thick] (m-2-10) -- (m-3-11);
    	\draw[red, -<] (a1) -- (m-3-3);
    	\draw[red, -<] (a1) -- (m-3-5);
    	\draw[red, ->] (a1) -- (m-2-4);
    	\draw[red, -<] (a2) -- (m-1-9);
    	\draw[red, densely dotted, thick] (a2) -- (m-1-11);
    	\draw[red, ->] (a2) -- (m-2-10);
    	\draw[red, -<] (b) -- (m-2-4);
    	\draw[red, -<] (b) -- (m-2-6);
    	\draw[red, ->] (b) -- (m-1-5);
    	\draw[red, ->] (b) -- (m-3-5);
    	\draw[red, -<] (c) -- (m-1-5);
    	\draw[red, -<] (c) -- (m-1-7);
    	\draw[red, ->] (c) -- (m-2-6);
    	\draw[red, -<] (d) -- (m-2-6);
    	\draw[red, -<] (d) -- (m-2-8);
    	\draw[red, ->] (d) -- (m-1-7);
    	\draw[red, ->] (d) -- (m-3-7);
    	\draw[red, densely dotted, thick] (e1) -- (m-1-1);
    	\draw[red, -<] (e1) -- (m-1-3);
    	\draw[red, ->] (e1) -- (m-2-2);
    	\draw[red, -<] (e2) -- (m-3-7);
    	\draw[red, -<] (e2) -- (m-3-9);
    	\draw[red, ->] (e2) -- (m-2-8);
    	\draw[red, -<] (f1) -- (m-2-2);
    	\draw[red, -<] (f1) -- (m-2-4);
    	\draw[red, ->] (f1) -- (m-1-3);
    	\draw[red, ->] (f1) -- (m-3-3);
    	\draw[red, -<] (f2) -- (m-2-8);
    	\draw[red, -<] (f2) -- (m-2-10);
    	\draw[red, ->] (f2) -- (m-1-9);
    	\draw[red, ->] (f2) -- (m-3-9);
    \end{tikzpicture}
    }
    % \vspace{-.1cm}
	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[\B_\circ] \big)$ for the cluster fan of \cref{fig:lodayFans}\,(right). See \cref{thm:extremalExchangeablePairsCA}. The conventions are similar to \cref{fig:labelFacetDefiningInequalititiesAsso}.}
	\label{fig:labelFacetDefiningInequalititiesCA}
\end{figure}

@@

\begin{theorem}
\label{thm:allPolytopalRealizationsCA}
For any finite type exchange matrix~$\B_\circ$, and for any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
\b{R}_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\variables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \!\!\!\!\!\! \sum_{y \in \variables} \!\!\!\!\!\coefficient[y][x][x'] \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$\b{R}_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{theorem}

@@

\item the non-kissing fan~$\gvectorFan[\quiver]$ is the normal fan of the \defn{gentle associahedron}~$\Asso[\quiver]$, constructed in~\cite{PaluPilaudPlamondon-nonkissing}.
